---
title: "Which country is the best sales destination for alcohol?"
author: "Kris"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---
```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

I was talking to a German friend one day about how good German are at drinking alcohol, he said "the Chinese moutai is nothing". While taste is kind of personal, I did not feel offended as we were drinking and telling nonsense as well. That was back in Denmark where student orientations typically have challenge activities such as drinking 6 litres of beer. Now, I am intrigued to find out how nationals of different countries love beer, spirits or wine. Perhaps this analysis will be a talking point of mine during interviews with Budwiser or Heineken. 

## Load packages

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
```


## Load Data

```{r, load_alcohol_data}
library(fivethirtyeight)
data(drinks)
```

## Take a general look at the data

```{r glimpse_skim_data}
glimpse(drinks)
skim(drinks)
```

Lucky! I have a great data set. Simply plot a bar graph. And the top countries must be chief markets for Carlsberg, Budweiser, Heineken etc.

## Countries by beer consumption

```{r beer_plot}
beer_plot <- drinks %>% 
  arrange(desc(beer_servings)) %>% 
  head(25)
ggplot(data = beer_plot, mapping = aes(reorder(x=country,beer_servings),y=beer_servings))+
  geom_col()+
  coord_flip()+
  labs(x="Country",y="Beer Servings", title="Top 25 Beer Consuming Countries")

```

I did not expect Namibia to come on top! And it is indeed my first time hearing about the country Gabon. But that is just life: realize the reality and change your mentality. Next time, when I visit West Africa, I for sure will get some beers and have fun with the locals.

## Countries by wine consumption

```{r wine_plot}
wine_plot <- drinks %>% 
  arrange(desc(wine_servings)) %>% 
  head(25)
ggplot(data = wine_plot, mapping = aes(reorder(x=country,wine_servings),y=wine_servings))+
  geom_col()+
  coord_flip()+
  labs(x="Country",y="Wine Servings",title= "Top 25 Wine Consuming Countries")
```

Interesting, the top 25 wine consuming countries are all European and Hispanic countries. Where is China? Let me propose a sales opportunity to producers at Bordeaux.

## Countries by spirits consumption

```{r spirit_plot}
spirit_plot <- drinks %>% 
  arrange(desc(spirit_servings)) %>% 
  head(25)
ggplot(data = beer_plot, mapping = aes(reorder(x=country,spirit_servings),y=spirit_servings))+
  geom_col()+
  coord_flip()+
  labs(x="Country",y="Spirit Servings",title="Top 25 Spirit Consuming Countries")
```

This one I kinda agree. Russians' been reputable for being a "fighting" race. Spirits, in a pun, sense are rooted in their bones... Look... No offense but I understand they need to warm themselves up boi!
