---
title: "Covid death rates by gender, co-morbidities, and ICU admission"
author: "Kris"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

The UK's 3rd nation-wide pandemics lockdown is scheduled 14th October. It will be very unfortunate to get infected the virus, especially that a cold winter is approaching. Let's look at the likelihood of death if you get them. The point is not to **scare** you, but to *warn* you to be careful!!!

```{r, setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

# Load data packages

```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
```

# Download and clean data frame

```{r, cache=TRUE}
# URL link to CDC to download data
url <- "https://data.cdc.gov/api/views/vbim-akqf/rows.csv?accessType=DOWNLOAD"

covid_data <- vroom::vroom(url)%>% # If vroom::vroom(url) doesn't work, use read_csv(url)
  clean_names()
```

# Graph of death rates by gender and co-morbidities

Are we more likely to die if we have co-morbidities?

1. by age group, sex, and whether the patient was admited to Intensive Care Unit (ICU) or not.

```{r,,,fig.width=12}
# Filter and summarize count, for grouped data, with death criteria
grouped_med_death <- covid_data %>%
  filter(medcond_yn %in% c("Yes", "No"),        # Get rid of "Unknown" or "Missing" medcond_yn
         death_yn %in% c("Yes", "No"),          # Get rid of "Missing" death_yn
         death_yn != "Unknown",                 # Get rid of "Unknown" death_yn
         sex %in% c("Male", "Female"),          # Get rid of sex other than "Male" and "Female"
         age_group != "Unknown") %>%            # Get rid of age_group "Unknown"
  group_by(age_group, sex, medcond_yn, death_yn) %>%           # Group, with "Yes" or "No" as separate categories for death_yn
  summarise(N=n())                              # Obtain count for combinations of categories

# Filter and summarize count, for grouped data, without death criteria
grouped_med <- covid_data %>%
  filter(medcond_yn %in% c("Yes", "No"),
         death_yn %in% c("Yes", "No"),
         death_yn != "Unknown",
         sex %in% c("Male", "Female"),
         age_group != "Unknown") %>%
  group_by(age_group, sex, medcond_yn) %>%                     # Group, regardless of "Yes" or "No" for death_yn
  summarise(TN=n())

# Join categorical count and total count, calculate death rate
plot_comorbidities <- grouped_med_death %>%
  left_join(grouped_med, by = c("age_group", "sex", "medcond_yn")) %>%         # Present total count next to sub-count
  mutate(death_rate = N / TN * 100,                                                    # Create new column of calculated percentage of total
         medcond_yn = factor(medcond_yn,                                               # Factorize and rename medcond_yn feature
                             levels = c("Yes", "No"),
                             labels = c("With comorbidities",
                                        "Without comorbidities"))) %>%
  filter(death_yn == "Yes")      # Only use death_rate (not non-death rate) before plotting

# Plot
ggplot(plot_comorbidities, aes(x = age_group, y = death_rate)) +               # Specify data and axes
  geom_col(fill = "#365383") +                                                    # Specify chart style (bar) and fill color
  geom_text(mapping = aes(y = death_rate + 3, label = round(death_rate,1)), size = 4) +  # Specify data labels and their positions
  coord_flip() +                                                               # Flip axis
  facet_grid(medcond_yn ~ sex) +                                               # Create a grid format (vertical ~ horizontal)
  labs(x="", 
       y = "", 
       title = "Covid death % by age group, sex and presence of co-morbidities") +
  scale_y_continuous(breaks = seq(0, 60, 20),
                     labels = paste(seq(0, 60, 20), "%")) +
  theme_bw()
```

Look, when you get comorbidities, do not be too optimistic of your health. You are more likely to *die* than to live!

# Graph of death rates by gender and icu admissions

Are we more likely to die if we are admitted to ICUs?

```{r,,,fig.width=12}
# Filter and summarize count, for grouped data, with death criteria
grouped_icu_death <- covid_data %>%
  filter(icu_yn %in% c("Yes", "No"),
         death_yn %in% c("Yes", "No"),
         sex %in% c("Male", "Female"),
         age_group != "Unknown") %>%
  group_by(age_group, sex, icu_yn, death_yn) %>%
  summarise(N=n())

# Filter and summarize count, for grouped data, without death criteria
grouped_icu <- covid_data %>%
  filter(icu_yn %in% c("Yes", "No"),
         death_yn %in% c("Yes", "No"),
         sex %in% c("Male", "Female"),
         age_group != "Unknown") %>%
  group_by(age_group, sex, icu_yn) %>%
  summarise(TN=n())

# Join categorical count and total count, calculate death rate
plot_icu <- grouped_icu_death %>%
  left_join(grouped_icu, by = c("age_group", "sex", "icu_yn")) %>%
  mutate(death_rate = N / TN * 100,
         icu_yn = factor(icu_yn, 
                         levels = c("Yes", "No"),
                         labels = c("Admited to ICU",
                                    "No ICU"))) %>%
  filter(death_yn == "Yes")

# Plot
ggplot(plot_icu, aes(x = age_group, y = death_rate)) +
  geom_col(fill = "#fc9272") +
  geom_text(mapping = aes(y = death_rate + 3, label = round(death_rate,1)), size = 4) +
  coord_flip() +
  facet_grid(icu_yn ~ sex) +
  labs(x="", 
       y = "", 
       title = "Covid death % by age group, sex and whether patient was admited to ICU") +
  scale_y_continuous(breaks = seq(0, 60, 20),
                     labels = paste(seq(0, 60, 20), "%")) +
  theme_bw()
```

Look, when you are admitted to an ICU, do not be too optimistic of your health. You are more likely to *die* than to live! Perhaps the logic is not that ICU treatment hastens your death. It is because you are dying that ICUs accepted you!

The point is, *mask yourself, keep distance* !

# Acknowledgement: [CDC Covid-19 Case Surveillance Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf). 

