---
title: "Return of Stocks"
author: "Kris"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

As a finance student, it is interesting to look at time series data and decide which index to invest. We shall see, for this page, how the NYSE-listed stocks performed.

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

## Load data packages

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
```


## Load data

```{r load_nyse_data, message=FALSE, warning=FALSE}
nyse <- read_csv(here::here("data","nyse.csv"))
```

First, it is a good practice to see what example industries the stocks belong to.

## Number of companies by sector 

```{r}
companies_per_sector <- nyse %>% 
  group_by(sector) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
companies_per_sector
```

## Visualize number of companies by sector 

```{r}
bar <- companies_per_sector %>% 
  ggplot(aes(y=reorder(sector,count),x=count))+
         geom_col()+
  labs(y="Sector", x="# of Companies",title="Number of Companies per Sector")
bar
```

For demonstration purpose, rather than analyzing company one by one, let's look at two indexes Dow Jones Industrials and S&P 500.

## Load DJIA and SPY data

```{r, tickers_from_wikipedia}

djia_url <- "https://en.wikipedia.org/wiki/Dow_Jones_Industrial_Average"

#get tables that exist on URL
tables <- djia_url %>% 
  read_html() %>% 
  html_nodes(css="table")


# parse HTML tables into a dataframe called djia. 
# Use purr::map() to create a list of all tables in URL
djia <- map(tables, . %>% 
               html_table(fill=TRUE)%>% 
               clean_names())


# constituents
table1 <- djia[[2]] %>% # the second table on the page contains the ticker symbols
  mutate(date_added = ymd(date_added),
         
         # if a stock is listed on NYSE, its symbol is, e.g., NYSE: MMM
         # We will get prices from yahoo finance which requires just the ticker
         
         # if symbol contains "NYSE*", the * being a wildcard
         # then we jsut drop the first 6 characters in that string
         ticker = ifelse(str_detect(symbol, "NYSE*"),
                          str_sub(symbol,7,11),
                          symbol)
         )

# we need a vector of strings with just the 30 tickers + SPY
tickers <- table1 %>% 
  select(ticker) %>% 
  pull() %>% # pull() gets them as a sting of characters
  c("SPY") # and let us add SPY, the SP500 ETF

```

## Arrange data and `glimpse`

```{r get_price_data, message=FALSE, warning=FALSE, cache=TRUE}
myStocks <- tickers %>% 
  tq_get(get  = "stock.prices",
         from = "2000-01-01",
         to   = "2020-08-31") %>%
  group_by(symbol) 

glimpse(myStocks) # examine the structure of the resulting data frame
```

## Calculate returns

```{r calculate_returns, message=FALSE, warning=FALSE, cache=TRUE}
#calculate daily returns
myStocks_returns_daily <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "daily", 
               type       = "log",
               col_rename = "daily_returns",
               cols = c(nested.col))  

#calculate monthly  returns
myStocks_returns_monthly <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "monthly_returns",
               cols = c(nested.col)) 

#calculate yearly returns
myStocks_returns_annual <- myStocks %>%
  group_by(symbol) %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "yearly", 
               type       = "arithmetic",
               col_rename = "yearly_returns",
               cols = c(nested.col))
```

## Compute stock return stats

```{r summarise_monthly_returns}

summarise_monthly_returns <- myStocks_returns_monthly %>% 
  summarise(min=min(monthly_returns),
            max=max(monthly_returns),
            median=median(monthly_returns),
            mean = mean(monthly_returns),
            SD=sd(monthly_returns))

summarise_monthly_returns
```

## Visualize distribution of returns

```{r density_monthly_returns}

density_monthly_returns <- myStocks_returns_monthly %>% 
  ggplot(aes(x=monthly_returns))+
  geom_density() + 
  facet_wrap(~symbol)

density_monthly_returns
```

Wow, even a myopic look could tell you that Dow has an unusually flat shape (lower kurtosis) compared to others. It seems therefore the riskiest as possible monthly return span out a further range, and also have many in the negative, with no indication of clear model points. S&P 500 is the least risky as movements of the returns are more concentrated than others.

# Return vs risk plot

```{r risk_return_plot}
risk_return_plot <- summarise_monthly_returns %>%
  ggplot(aes(x=SD,y=mean))+
  ggrepel::geom_text_repel(aes(label = symbol))+
  labs(x="Risk", y= "Expected Monthly Return")

risk_return_plot
```

It looks like DOW is the riskiest because it has the least mean/standard deviation ratio, suggesting that the index awards the least return per unit risk born.