---
title: "Fuel efficiency and engine"
author: "Kris"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: zenburn
    theme: flatly
    toc: yes
    toc_float: yes
---

Although intuitive, one may conduct a statistical analysis on how engine size impact fuel efficiency. In order to do that, we must first call out the packages.

# Load packages
```{r load-libraries, warning=FALSE, message=FALSE}
# Load ggplot2, dplyr, and all the other tidyverse packages
library(tidyverse)  
library(mosaic)
library(lubridate)
library(fivethirtyeight)
```

# Glimpse Dataframe
Next, it is good practice to take a look at the dataframe mpg using a `glimpse` function

```{r}
# Have a look at how the dataframe is structured
glimpse(mpg)
```

We can observe that the dataframe has 234 rows and 11 columns.
- `manufacturer` specifies the brand.
- `model` specifies the model name under the brand
- `displ` means engine displacement, in litres
- `year` is the year of manufacture of the model
- `cyl` records the number of cylinders of the engine
- `trans` specifies type of transmission
- `drv` means the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd
- `cty` records the driving fuel efficiency in cities: city miles per gallon
- `hwy` records the driving fuel efficiency on highways: highway miles per gallon
- `fl` is fuel type
- `class` is "type" of car

# Visualize relationship between highway fuel efficiency and number of cylinder 
If you liked the "Need for Speed" feeling of fastness, you probably want more cylinders in the car engine to power a hasty start. What if you are yet environmentally conscious and want a good fuel efficiency as well? Is there a trade-off?

```{r}
# Make a scatterplot of hwy vs cyl
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy))
```

Clearly, there is a trade-off! It looks like hastier accelerations with more cylinders cost more fuels!

# Visualize relationship between highway fuel efficiency and displacement
A better approximation of engine power, rather than cylinder number, is engine displacement, i.e. how much fuel gets absorbed into the engine in one round of piston motion. How may fuel efficiency be correlated with engine displacement? Thus, if we want to check our carbon footprint, do we still want to buy cars with immense fuel displacement? Does the type of car matter in our analysis?

```{r}
# Make a scatterplot of hwy vs cyl
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

Hooyay! It looks like compact cars have the highest fuel efficiency. No wonder you see them clouding around in Europe! And SUVs are probably the least efficient behemoths. But if you like to conquer hills, muds, and prairies, have them around. 2 seaters are sports cars. Looks like they are the outliers at high range of displacement. But, consider your pocket and road conditions before purchase!

```{r, echo=FALSE}
 # Read [Chapter 3 of R for Data Science](http://r4ds.had.co.nz/data-visualisation.html) and complete the following exercises:

# [3.2.4](http://r4ds.had.co.nz/data-visualisation.html#exercises): Questions 1–5
# [3.3.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-1): Questions 1–5
# [3.5.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-2): Questions 1–4
# [3.6.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-3): Questions 1–5
# [3.8.1](http://r4ds.had.co.nz/data-visualisation.html#exercises-5): Questions 1 and 2
```



